---
title: "xtranat"
subtitle: "Application of Network Analysis Techniques to Input-Output Economic Systems"
author:
- name: "Fernando DePaolis(*)"
  affiliation: . [Middlebury Institute of International Studies](http://www.miis.edu/) [CBE]( http://centerfortheblueeconomy.org/)
- name: "Phil Murphy"
  affiliation: . [Middlebury Institute of International Studies](http://www.miis.edu/) [META Lab](http://sites.miis.edu/metalab/)
- name: "M. Clara DePaolis Kaluza"
  affiliation: Northeastern University, College of Computer and Information Science
date: "March, 2018"
output:
  html_document:
    fig_height: 6
    theme: spacelab
    toc: yes
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{xtranat}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
abstract: "These script chunks are part of the 'xtranat' package. They compute 'Counting Betweenness' and 'Random Walk Centrality' from Input-Output data from IMPLAN(R). It also computes the intermediate measure 'Mean First Pass Time'. The scripts have been developed for R from the MatLab code in Blöchl et al (2011)."
keywords: "Input-Output system; Regional Economies; Multiplier effects"

geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
---

<div style="text-align: right;>(*) Corresponding author: Fernando DePaolis [fdepaolis@middlebury.edu]</div>

#### Mean First Pass Time
```{r}
mfpt <- function(A) {
        A <- as.matrix(A)  ## IO coeffs, or dollar values?
        n = nrow(A) ## total number of sectors (20, 86 or 536)
        rrss = rowSums(A)
        for (i in 1:n) {
                if (rrss[i] != 0) {
                        rrss[i] = 1/rrss[i]
                }
        }
        AA = diag(n) - diag(rrss) %*% A  #compute transition matrix. Is this "T" in 'mediative effects' Garcia-Muniz 2008?
        H = matrix(0, n, n)
        I = solve(AA[-1,-1])            ## inverse of AA without 1st column & 1st row
        ones = matrix(1, n-1, 1)        ## vector of "1"s of length 'n-1'
        for (i in 1:n) {
                H[-i,i] = I %*% ones    ## matrix product; otherwise, non-conformable
                if (i < n){
                        u = AA[-(i+1),i] - AA[-i, (i+1)]
                        I = I - ((I*u) * I[i,]) / (1 + (I[i,] * u))
                        v = AA[i, -(i+1)] - AA[(i+1), -i]
                        I = I - ((I[,i] * (v * I)) / 1 + v * I[,i])
                        if (AA[(i+1),(i+1)]!=1){
                                I = solve(AA[-(i+1),-(i+1)], tol = 1e-29)
                        }
                        if (any(is.infinite(I))) {        ## i.e. Sherman-Morrison didn't work. When would I(i,j)=infinity
                                I[is.infinite(I)] <- 0
                        }
                }
        }
        H <<- H
}

```
<div style="text-align: right;" markdown="1"><a href="#header">back to top</a></div>

#### Counting Betweenness
```{r}
cbet <- function(A) {
        ## Reads the A-matrix; removes row/column with zeros; records their row/column number
        A <- as.matrix(A)
        m = nrow(A)
        rrss = rowSums(A)
        retain.vector <- vector(mode="numeric", length=0)
        if (0.0 %in% rrss){    ## Checks if there is a row with all zeros
          retain.vector <- row(as.matrix(rrss))[which(as.matrix(rrss) == 0)]
          AA1 = A[-retain.vector,-retain.vector]  ## this is the A-matrix without row/columns of zeros
        } else {
          AA1 = A
        }

        d = diag(rowSums(AA1))
        n = nrow(AA1)
        ones = matrix(1, n, 1) ## this is a vector of "n" rows by 1 col of "1"
        re = matrix(0, n, 1 )  ## this is a vector of "n" rows by 1 col of "0"
        for (p in 1:n){
                atemp = AA1[-p,-p]
                T = solve(d[-p,-p] - atemp, tol = 1e-29)
                for (s in 1:n){
                        if (s != p){
                                if (s < p){
                                    indx = s
                                } else if (s > p) {
                                    indx = s - 1
                                }
                        N = as.matrix(diag(T[indx,])) %*% atemp
                        I = abs(N + t(N)) / 2
                        re[-p,1] = re[-p,1] + 0.5*((t(colSums(I))) + rowSums(I))
                        }
                }
        }
        re2 = (re + 2 * (n-1) * ones) / ((n) * (n-1))

        res = matrix(0, m, 1)
        # restore one or more rows/columns of zeros to their original positions
        if (length(retain.vector)!=0) {
          res[-retain.vector] <- re2
        } else
          res <- re2

        res <<- res
}
````
<div style="text-align: right;" markdown="1"><a href="#header">back to top</a></div>

#### Random Walk Centrality
```{r}
rwc <- function(A) {
        nn = nrow(A)
        cen = matrix(0,nn,1)
        m <- mfpt(A)   # H from mfpt{}
        for (j in 1:nn) {
                if (all(H[j,] == (c(rep(1,(j-1)),0,rep(1,(nn-j)))))) {  # This compares each row of H with a rows made of 1s and a zero on the diagonal
                        cen[j] = 0   # If TRUE (i.e. row of H == 1s) that row of CEN == zero
                } else {
                        cen[j] = nn / sum(m[,j])
                }
        }
        cen <<- cen
}
````
<div style="text-align: right;" markdown="1"><a href="#header">back to top</a></div>

### Closing Comments
Based on Blöchl F, Theis FJ, Vega-Redondo F, and Fisher E: Vertex Centralities in Input-Output Networks Reveal the Structure of Modern Economies, Physical Review E, 83(4):046127, 2011.[Link](https://journals.aps.org/pre/abstract/10.1103/PhysRevE.83.046127)

<div style="text-align: right;" markdown="1"><a href="#header">back to top</a></div>
